### 2024-09-26 paper/paper.qmd
wrote more in the paper

---
title: "Temporal and Spatial Analysis of Hate Crimes in Toronto"
subtitle: "Uncovering Neighborhood Disparities and Bias Trends Across the City"
author: 
  - Tommy Fu
thanks: "Code and data are available at: https://github.com/YichengFu/hate_crimes.git"
date: 24 September 2024
date-format: long
abstract: "This paper analyzes the temporal and spatial patterns of hate crimes in Toronto focusing on trends over time and neighborhood disparities. Toronto Open Data hate crime is used to explore variations in crime frequency and the prevalence of biases such as race, religion and sexual orientation. The results highlight distinct trends and reveal neighborhoods with higher concentrations of bias-motivated incidents; these findings offer insights for addressing hate crimes in the city."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(dplyr)
library(knitr)
library(ggplot2)
library(readr)
library(tinytex)

#load the data
analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
```

# Introduction

Hate crimes are a significant incident reflecting deep-rooted prejudice and discrimination within communities. Hate-crime victimization against racially visible people is of growing concern (Chongatera 2013). These crimes not only impact individuals but also harm the society often leaving individuals feeling unsafe. Social violence omit both the daily violence suffered by certain social categories and its many impacts on the victims (Dalphond 2021). In Toronto, understanding the patterns and dynamics of hate crimes is crucial for addressing their causes and mitigating their effects. While racial and cultural diversity initiatives are central in hate crime policy, combating racially motivated hate crime is often obscured by matters considered more significant by police (Bryan 2019). This paper aims to fill that gap by providing an in-depth analysis of hate crimes in Toronto over time and across different neighborhoods.

This study focuses exploring two primary questions: how hate crimes in Toronto have evolved over time and whether certain neighborhoods experience a disproportionate concentration of bias-motivated incidents. Using Toronto open Data hate crime from 2018 to 2023 I analyzes temporal trends to identify peaks and patterns in the frequency of reported incidents. Additionally, the research examines whether specific biases (such as race, religion, or sexual orientation) are more prevalent in particular areas contributing to an understanding of neighborhood-level disparities. The results provide insight into both the temporal and spatial aspects of hate crimes, shedding light on how bias manifests in different contexts within the city.

The data section will introduced the detail of the data set used in this research in @sec-data -- Data. The variables as well as the cleaning process will be discussed in this section. @sec-results will focus on the findings and the visualization of the analysis. @sec-discussion talks about the limitation of this research and potential error caused by research design or the natural form of data. lastly, the conclusion part raps up the discovers and summarizes all the findings into a short paragraph.

# Data {#sec-data}

## Measurement and packages 
The Data set used in the analysis is gathered from Open Data Toronto through the Open Data Toronto [@opendatatoronto] and used the statistical software R [@citeR] for importing data, data cleaning and testing data. The data source "Crime Data" is collected from Toronto Police starting from 2018 to the end of 2023. It includes both temporal and spatial attributes. Other packages were used for analyzing the data such as ggplot [@ggplot2], knitr[@knitr], tidyverse [@tidyverse], lubridate [@lubridate], tinytex[@tinytex], dplyr [@dplyr].

Data contains around 1400 observations and contains the occurrence date and specific time where the incident happened and the date reported to the police. The reason for the hate crime is categorized as race bias, religion bias and sexual orientation bias etc. The motive for 
committing a crime, the illegal activity and the neighborhood where the crime occurred are documented by the police throughout the years. Furthermore, the type of location  such as park, apartment and school are documented to form the detailed data set and is named "Hate Crime".

## Data cleaning

After observing the data, some observations were dropped due to missing values in location type meaning the recorder had a hard time describing the location characteristics. Therefore, these observations were dropped out of the cleaned data set and will not be included in further data analysis. Date is critical in our data investigation, observations where the reported time is before occurrence time is dropped to assure data validity. Since the data set is from Open Data Toronto the quality of the data is decent. Further cleaning process is unnecessary due to origin high quality of data set.

## Observation of cleaned data
A sample of observation of cleaned data is displayed in @tbl-1. This table includes key variables of interest that will be used for further analysis. Specifically, it contains the bias type, which indicates whether the crime was motivated by factors such as race, religion, or sexual orientation. The date of the incident is provided, allowing for temporal analysis of hate crime trends over time. Additionally, the neighborhood where each crime occurred is included, which is crucial for understanding geographic characteristics and identifying potential hotspots for hate crimes within Toronto. The offense type records the legal classification of the crime, such as assault or mischief providing insight into the nature of the incidents. Lastly, the arrest status variable indicates whether law enforcement made an arrest in response to the reported crime, which is an important indicator of justice system responses to hate crimes.

```{r}
#| label: tbl-1
#| fig-cap: Sample data visualization of what the data looks like containing the variable of intersts. Note the column sex. stands for sexual orientation. RLGbias stands for religion bias.
#| echo: false
#| warning: false
#| message: false

# Rename columns for a shorter format
selected_variables <- analysis_data %>%
  select(OCCURRENCE_YEAR, RACE_BIAS,  SEXUAL_ORIENTATION_BIAS, RELIGION_BIAS,
         NEIGHBOURHOOD_158, PRIMARY_OFFENCE, ARREST_MADE) %>%
  rename(
    Year = OCCURRENCE_YEAR,
    RaceBias = RACE_BIAS,
    Sex. = SEXUAL_ORIENTATION_BIAS,
    RLGBias = RELIGION_BIAS,
    Location = NEIGHBOURHOOD_158,
    Offense = PRIMARY_OFFENCE,
    Arrest = ARREST_MADE
  )

# Display a sample of 8 rows from the renamed variables
sample_data <- selected_variables %>%
  sample_n(8)

# Show the table using kable
kable(sample_data, caption = "Sample of Cleaned Data with Selected Variables", align = "c")
```
## Basic Summary Statistics of the Data

A general summary statistic is shown @tbl-2. Since hate crimes are relatively rare events, the data may show an uneven distribution across different years and neighborhoods. Some years may exhibit higher counts due to specific incidents or socio-political factors that led to spikes in reporting, while other years may show fewer cases, resulting in an unbalanced distribution. Due to the categorical nature of many key variables, such as bias types (race, religion, sexual orientation), and the relatively small number of observations per category, more advanced summary statistics (e.g., mean, standard deviation) are not as meaningful in this context. Instead, the focus is on counting occurrences and understanding the distribution of incidents across time and location. Therefore, we limit the analysis to counts and proportions rather than providing more traditional numeric summary statistics. Further analysis of location and its relationship to number of hate crimes will be shown in @sec-results.

```{r}
#| label: tbl-2
#| echo: false
#| warning: false
#| message: false
# Count the number of observations (hate crimes) by year
observations_by_year <- analysis_data %>%
  group_by(OCCURRENCE_YEAR) %>%
  summarise(Observation_Count = n()) %>%
  arrange(OCCURRENCE_YEAR)

# Display the table with kable
kable(observations_by_year, caption = "Number of Observations by Year", align = "c")

```

# Results {#sec-results}

A sample observation is shown in @tbl-3. This table indicates neighborhoods in Toronto with more than 20 Hate Crimes from 2018 to 2023. The number of 20 hate crimes is manually selected by the author after inspecting the data to avoid overlapping of the visualization of the table or cluster of data. Only the significant neighbourhoods are selected to include in this table to give a brief observation of the number of crime case numbers in different area in Toronto. 
```{r}
#| label: tbl-3
#| tbl-cap: Neighborhoods with more than 20 Hate Crimes from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Filter for neighborhoods with more than 20 hate crimes
neighborhood_filtered <- analysis_data %>%
  group_by(NEIGHBOURHOOD_158) %>%
  summarise(Crime_Count = n()) %>%
  filter(Crime_Count > 20) %>%
  arrange(desc(Crime_Count))

# Display the table using kable.
kable(neighborhood_filtered)
```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References